\bf Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-03-27 paper.qmd
Added figure description

@@

umap_plot_pbmc <- umap_pbmc_scaled |>
  ggplot(aes(x = UMAP1,
             y = UMAP2))+
  geom_point(alpha=0.5) +
  coord_equal() +
  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + 
  theme_linedraw() +
  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
        legend.title = element_text(size=5), #change legend title font size
        legend.text = element_text(size=4),
        legend.key.height = unit(0.25, 'cm'),
        legend.key.width = unit(0.25, 'cm')) 
```

```{r}
#| echo: false
#| fig-cap: "2D layout from UMAP (n_neighbors = 30, min_dist = 0.3) applied for the PBMC3k dataset. Is this a best representation of the original data?"
#| label: fig-umapauthor
#| fig-pos: H

umap_plot_pbmc 
```

>>>>

## Introduction

@@

umap_plot_pbmc <- umap_pbmc_scaled |>
  ggplot(aes(x = UMAP1,
             y = UMAP2))+
  geom_point(alpha=0.1, size=1) +
  coord_equal() +
  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + 
  theme_linedraw() +
  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
        legend.title = element_text(size=5), #change legend title font size
        legend.text = element_text(size=4),
        legend.key.height = unit(0.25, 'cm'),
        legend.key.width = unit(0.25, 'cm')) +
  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3) 
```

```{r}
#| warning: false
#| echo: false

umap_pbmc <- read_rds("data/pbmc3k/pbmc_umap_5_min_dist_0.01.rds")
umap_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = umap_pbmc, x = "UMAP1", y = "UMAP2"))) |>
  dplyr::rename(c("UMAP1" = "scaled_UMAP1", 
                  "UMAP2" = "scaled_UMAP2")) |>
  dplyr::mutate(ID = 1:NROW(umap_pbmc))

umap_plot_pbmc_1 <- umap_pbmc_scaled |>
  ggplot(aes(x = UMAP1,
             y = UMAP2))+
  geom_point(alpha=0.1, size=1) +
  coord_equal() +
  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + 
  theme_linedraw() +
  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
        legend.title = element_text(size=5), #change legend title font size
        legend.text = element_text(size=4),
        legend.key.height = unit(0.25, 'cm'),
        legend.key.width = unit(0.25, 'cm')) +
  annotate(geom = 'text', label = 'b', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)

```

```{r}
#| warning: false
#| echo: false

umap_pbmc <- read_rds("data/pbmc3k/pbmc_umap_15_min_dist_0.99.rds")
umap_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = umap_pbmc, x = "UMAP1", y = "UMAP2"))) |>
  dplyr::rename(c("UMAP1" = "scaled_UMAP1", 
                  "UMAP2" = "scaled_UMAP2")) |>
  dplyr::mutate(ID = 1:NROW(umap_pbmc))

umap_plot_pbmc_2 <- umap_pbmc_scaled |>
  ggplot(aes(x = UMAP1,
             y = UMAP2))+
  geom_point(alpha=0.1, size=1) +
  coord_equal() +
  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + 
  theme_linedraw() +
  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
        legend.title = element_text(size=5), #change legend title font size
        legend.text = element_text(size=4),
        legend.key.height = unit(0.25, 'cm'),
        legend.key.width = unit(0.25, 'cm')) +
  annotate(geom = 'text', label = 'c', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

```

```{r}
#| warning: false
#| echo: false

umap_pbmc <- read_rds("data/pbmc3k/pbmc_umap_15_min_dist_0.03.rds")
umap_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = umap_pbmc, x = "UMAP1", y = "UMAP2"))) |>
  dplyr::rename(c("UMAP1" = "scaled_UMAP1", 
                  "UMAP2" = "scaled_UMAP2")) |>
  dplyr::mutate(ID = 1:NROW(umap_pbmc))

umap_plot_pbmc_3 <- umap_pbmc_scaled |>
  ggplot(aes(x = UMAP1,
             y = UMAP2))+
  geom_point(alpha=0.1, size=1) +
  coord_equal() +
  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + 
  theme_linedraw() +
  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
        legend.title = element_text(size=5), #change legend title font size
        legend.text = element_text(size=4),
        legend.key.height = unit(0.25, 'cm'),
        legend.key.width = unit(0.25, 'cm')) +
  annotate(geom = 'text', label = 'd', x = Inf, y = Inf, hjust = 1.5, vjust = 1.3, size = 3)

```

```{r}
#| warning: false
#| echo: false

umap_pbmc <- read_rds("data/pbmc3k/pbmc_umap_84_min_dist_0.99.rds")
umap_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = umap_pbmc, x = "UMAP1", y = "UMAP2"))) |>
  dplyr::rename(c("UMAP1" = "scaled_UMAP1", 
                  "UMAP2" = "scaled_UMAP2")) |>
  dplyr::mutate(ID = 1:NROW(umap_pbmc))

umap_plot_pbmc_4 <- umap_pbmc_scaled |>
  ggplot(aes(x = UMAP1,
             y = UMAP2))+
  geom_point(alpha=0.1, size=1) +
  coord_equal() +
  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + 
  theme_linedraw() +
  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
        legend.title = element_text(size=5), #change legend title font size
        legend.text = element_text(size=4),
        legend.key.height = unit(0.25, 'cm'),
        legend.key.width = unit(0.25, 'cm')) +
  annotate(geom = 'text', label = 'e', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)

```

```{r}
#| warning: false
#| echo: false
tsne_pbmc <- read_rds("data/pbmc3k/pbmc_tsne_5.rds")
tsne_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = tsne_pbmc, 
                                                                 x = "tSNE1", y = "tSNE2"))) |>
  dplyr::rename(c("tSNE1" = "scaled_tSNE1", 
                  "tSNE2" = "scaled_tSNE2")) |>
  dplyr::mutate(ID = 1:NROW(tsne_pbmc))

tsne_plot_pbmc_1 <- tsne_pbmc_scaled |>
  ggplot(aes(x = tSNE1,
             y = tSNE2))+
  geom_point(alpha=0.1, size=1) +
  coord_equal() +
  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
  theme_linedraw() +
  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
        legend.title = element_text(size=5), #change legend title font size
        legend.text = element_text(size=4),
        legend.key.height = unit(0.25, 'cm'),
        legend.key.width = unit(0.25, 'cm')) +
  annotate(geom = 'text', label = 'f', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)   
```

```{r}
#| warning: false
#| echo: false
tsne_pbmc <- read_rds("data/pbmc3k/pbmc_tsne_51.rds")
tsne_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = tsne_pbmc, 
                                                                 x = "tSNE1", y = "tSNE2"))) |>
  dplyr::rename(c("tSNE1" = "scaled_tSNE1", 
                  "tSNE2" = "scaled_tSNE2")) |>
  dplyr::mutate(ID = 1:NROW(tsne_pbmc))

tsne_plot_pbmc <- tsne_pbmc_scaled |>
  ggplot(aes(x = tSNE1,
             y = tSNE2))+
  geom_point(alpha=0.1, size=1) +
  coord_equal() +
  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
  theme_linedraw() +
  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
        legend.title = element_text(size=5), #change legend title font size
        legend.text = element_text(size=4),
        legend.key.height = unit(0.25, 'cm'),
        legend.key.width = unit(0.25, 'cm')) +
  annotate(geom = 'text', label = 'g', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.1, size = 3)   
```

```{r}
#| warning: false
#| echo: false
tsne_pbmc <- read_rds("data/pbmc3k/pbmc_tsne_92.rds")
tsne_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = tsne_pbmc, 
                                                                 x = "tSNE1", y = "tSNE2"))) |>
  dplyr::rename(c("tSNE1" = "scaled_tSNE1", 
                  "tSNE2" = "scaled_tSNE2")) |>
  dplyr::mutate(ID = 1:NROW(tsne_pbmc))

tsne_plot_pbmc_2 <- tsne_pbmc_scaled |>
  ggplot(aes(x = tSNE1,
             y = tSNE2))+
  geom_point(alpha=0.1, size=1) +
  coord_equal() +
  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
  theme_linedraw() +
  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
        legend.title = element_text(size=5), #change legend title font size
        legend.text = element_text(size=4),
        legend.key.height = unit(0.25, 'cm'),
        legend.key.width = unit(0.25, 'cm')) +
  annotate(geom = 'text', label = 'h', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)   
```

```{r}
#| warning: false
#| echo: false

phate_pbmc <- read_rds("data/pbmc3k/pbmc_phate_2.rds")
phate_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = phate_pbmc, x = "PHATE1", y = "PHATE2"))) |>
  dplyr::rename(c("PHATE1" = "scaled_PHATE1", 
                  "PHATE2" = "scaled_PHATE2")) |>
  dplyr::mutate(ID = 1:NROW(phate_pbmc))

phate_plot_pbmc_1 <- phate_pbmc_scaled |>
  ggplot(aes(x = PHATE1,
             y = PHATE2))+
  geom_point(alpha=0.1, size=1) +
  coord_equal() +
  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + 
  theme_linedraw() +
  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
        legend.title = element_text(size=5), #change legend title font size
        legend.text = element_text(size=4),
        legend.key.height = unit(0.25, 'cm'),
        legend.key.width = unit(0.25, 'cm')) +
  annotate(geom = 'text', label = 'i', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)

```

```{r}
#| warning: false
#| echo: false

phate_pbmc <- read_rds("data/pbmc3k/pbmc_phate_5.rds")
phate_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = phate_pbmc, x = "PHATE1", y = "PHATE2"))) |>
  dplyr::rename(c("PHATE1" = "scaled_PHATE1", 
                  "PHATE2" = "scaled_PHATE2")) |>
  dplyr::mutate(ID = 1:NROW(phate_pbmc))

phate_plot_pbmc_2 <- phate_pbmc_scaled |>
  ggplot(aes(x = PHATE1,
             y = PHATE2))+
  geom_point(alpha=0.1, size=1) +
  coord_equal() +
  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + 
  theme_linedraw() +
  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
        legend.title = element_text(size=5), #change legend title font size
        legend.text = element_text(size=4),
        legend.key.height = unit(0.25, 'cm'),
        legend.key.width = unit(0.25, 'cm')) +
  annotate(geom = 'text', label = 'j', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)

```

```{r}
#| warning: false
#| echo: false

phate_pbmc <- read_rds("data/pbmc3k/pbmc_phate_10.rds")
phate_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = phate_pbmc, x = "PHATE1", y = "PHATE2"))) |>
  dplyr::rename(c("PHATE1" = "scaled_PHATE1", 
                  "PHATE2" = "scaled_PHATE2")) |>
  dplyr::mutate(ID = 1:NROW(phate_pbmc))

phate_plot_pbmc_3 <- phate_pbmc_scaled |>
  ggplot(aes(x = PHATE1,
             y = PHATE2))+
  geom_point(alpha=0.1, size=1) +
  coord_equal() +
  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + 
  theme_linedraw() +
  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
        legend.title = element_text(size=5), #change legend title font size
        legend.text = element_text(size=4),
        legend.key.height = unit(0.25, 'cm'),
        legend.key.width = unit(0.25, 'cm')) +
  annotate(geom = 'text', label = 'k', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

```

```{r}
#| warning: false
#| echo: false

trimap_pbmc <- read_rds("data/pbmc3k/pbmc_trimap_12_4_3.rds")
trimap_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = trimap_pbmc, x = "TriMAP1", y = "TriMAP2"))) |>
  dplyr::rename(c("TriMAP1" = "scaled_TriMAP1",
                  "TriMAP2" = "scaled_TriMAP2")) |>
  dplyr::mutate(ID = 1:NROW(trimap_pbmc))

trimap_plot_pbmc_1 <- trimap_pbmc_scaled |>
  ggplot(aes(x = TriMAP1,
             y = TriMAP2))+
  geom_point(alpha=0.1, size=1) +
  coord_equal() +
  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) +
  theme_linedraw() +
  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
        legend.title = element_text(size=5), #change legend title font size
        legend.text = element_text(size=4),
        legend.key.height = unit(0.25, 'cm'),
        legend.key.width = unit(0.25, 'cm')) +
  annotate(geom = 'text', label = 'l', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)

```

```{r}
#| warning: false
#| echo: false

trimap_pbmc <- read_rds("data/pbmc3k/pbmc_trimap_30_4_10.rds")
trimap_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = trimap_pbmc, x = "TriMAP1", y = "TriMAP2"))) |>
  dplyr::rename(c("TriMAP1" = "scaled_TriMAP1",
                  "TriMAP2" = "scaled_TriMAP2")) |>
  dplyr::mutate(ID = 1:NROW(trimap_pbmc))

trimap_plot_pbmc_2 <- trimap_pbmc_scaled |>
  ggplot(aes(x = TriMAP1,
             y = TriMAP2))+
  geom_point(alpha=0.1, size=1) +
  coord_equal() +
  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) +
  theme_linedraw() +
  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
        legend.title = element_text(size=5), #change legend title font size
        legend.text = element_text(size=4),
        legend.key.height = unit(0.25, 'cm'),
        legend.key.width = unit(0.25, 'cm')) +
  annotate(geom = 'text', label = 'm', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)

```

```{r}
#| warning: false
#| echo: false

trimap_pbmc <- read_rds("data/pbmc3k/pbmc_trimap_5_2_2.rds")
trimap_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = trimap_pbmc, x = "TriMAP1", y = "TriMAP2"))) |>
  dplyr::rename(c("TriMAP1" = "scaled_TriMAP1",
                  "TriMAP2" = "scaled_TriMAP2")) |>
  dplyr::mutate(ID = 1:NROW(trimap_pbmc))

trimap_plot_pbmc_3 <- trimap_pbmc_scaled |>
  ggplot(aes(x = TriMAP1,
             y = TriMAP2))+
  geom_point(alpha=0.1, size=1) +
  coord_equal() +
  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) +
  theme_linedraw() +
  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
        legend.title = element_text(size=5), #change legend title font size
        legend.text = element_text(size=4),
        legend.key.height = unit(0.25, 'cm'),
        legend.key.width = unit(0.25, 'cm')) +
  annotate(geom = 'text', label = 'n', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

```

```{r}
#| warning: false
#| echo: false

pacmap_pbmc <- read_rds("data/pbmc3k/pbmc_pacmap_10_pca_0.5_2.rds")
pacmap_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = pacmap_pbmc, x = "PaCMAP1", y = "PaCMAP2"))) |>
  dplyr::rename(c("PaCMAP1" = "scaled_PaCMAP1",
                  "PaCMAP2" = "scaled_PaCMAP2")) |>
  dplyr::mutate(ID = 1:NROW(pacmap_pbmc))

pacmap_plot_pbmc_1 <- pacmap_pbmc_scaled |>
  ggplot(aes(x = PaCMAP1,
             y = PaCMAP2))+
  geom_point(alpha=0.1, size=1) +
  coord_equal() +
  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) +
  theme_linedraw() +
  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
        legend.title = element_text(size=5), #change legend title font size
        legend.text = element_text(size=4),
        legend.key.height = unit(0.25, 'cm'),
        legend.key.width = unit(0.25, 'cm')) +
  annotate(geom = 'text', label = 'p', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

```

```{r}
#| warning: false
#| echo: false

pacmap_pbmc <- read_rds("data/pbmc3k/pbmc_pacmap_30_random_0.9_5.rds")
pacmap_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = pacmap_pbmc, x = "PaCMAP1", y = "PaCMAP2"))) |>
  dplyr::rename(c("PaCMAP1" = "scaled_PaCMAP1",
                  "PaCMAP2" = "scaled_PaCMAP2")) |>
  dplyr::mutate(ID = 1:NROW(pacmap_pbmc))

pacmap_plot_pbmc_2 <- pacmap_pbmc_scaled |>
  ggplot(aes(x = PaCMAP1,
             y = PaCMAP2))+
  geom_point(alpha=0.1, size=1) +
  coord_equal() +
  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) +
  theme_linedraw() +
  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
        legend.title = element_text(size=5), #change legend title font size
        legend.text = element_text(size=4),
        legend.key.height = unit(0.25, 'cm'),
        legend.key.width = unit(0.25, 'cm')) +
  annotate(geom = 'text', label = 'q', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

```

```{r}
#| warning: false
#| echo: false

pacmap_pbmc <- read_rds("data/pbmc3k/pbmc_pacmap_5_pca_0.1_1.rds")
pacmap_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = pacmap_pbmc, x = "PaCMAP1", y = "PaCMAP2"))) |>
  dplyr::rename(c("PaCMAP1" = "scaled_PaCMAP1",
                  "PaCMAP2" = "scaled_PaCMAP2")) |>
  dplyr::mutate(ID = 1:NROW(pacmap_pbmc))

pacmap_plot_pbmc_3 <- pacmap_pbmc_scaled |>
  ggplot(aes(x = PaCMAP1,
             y = PaCMAP2))+
  geom_point(alpha=0.1, size=1) +
  coord_equal() +
  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) +
  theme_linedraw() +
  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
        legend.title = element_text(size=5), #change legend title font size
        legend.text = element_text(size=4),
        legend.key.height = unit(0.25, 'cm'),
        legend.key.width = unit(0.25, 'cm')) +
  annotate(geom = 'text', label = 'r', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

```

```{r}
#| warning: false
#| echo: false

pacmap_pbmc <- read_rds("data/pbmc3k/pbmc_pacmap_30_random_0.1_1.rds")
pacmap_pbmc_scaled <- as.data.frame(do.call(cbind, gen_scaled_data(data = pacmap_pbmc, x = "PaCMAP1", y = "PaCMAP2"))) |>
  dplyr::rename(c("PaCMAP1" = "scaled_PaCMAP1",
                  "PaCMAP2" = "scaled_PaCMAP2")) |>
  dplyr::mutate(ID = 1:NROW(pacmap_pbmc))

pacmap_plot_pbmc_4 <- pacmap_pbmc_scaled |>
  ggplot(aes(x = PaCMAP1,
             y = PaCMAP2))+
  geom_point(alpha=0.1, size=1) +
  coord_equal() +
  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) +
  theme_linedraw() +
  theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
        legend.title = element_text(size=5), #change legend title font size
        legend.text = element_text(size=4),
        legend.key.height = unit(0.25, 'cm'),
        legend.key.width = unit(0.25, 'cm')) +
  annotate(geom = 'text', label = 's', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)

```

```{r}
#| echo: false
#| fig-cap: "2D layouts from different NLDR techniques and different hyperparameter choices applied for the PBMC3k dataset: (a) UMAP (n_neighbors = 30, min_dist = 0.3), (b) UMAP (n_neighbors = 5, min_dist = 0.01), (c) UMAP (n_neighbors = 15, min_dist = 0.99), (d) UMAP (n_neighbors = 15, min_dist = 0.03), (e) UMAP (n_neighbors = 84, min_dist = 0.99), (f) tSNE (perplexity = 5), (g) tSNE (perplexity = 51), (h) tSNE (perplexity = 92), (i) PHATE (knn = 2), (j) PHATE (knn = 5), (k) PHATE (knn = 10), (l) TriMAP (n_inliers = 12, n_outliers = 4, n_random = 3), (m) TriMAP (n_inliers = 30, n_outliers = 4, n_random = 10), (n) TriMAP (n_inliers = 5, n_outliers = 2, n_random = 2), (p) PaCMAP (n_neighbors = 10, init = pca, MN_ratio = 0.5, FP_ratio = 2), (q) PaCMAP (n_neighbors = 30, init = random, MN_ratio = 0.9, FP_ratio = 5), (r) PaCMAP (n_neighbors = 5, init = pca, MN_ratio = 0.1, FP_ratio = 1), and (s) PaCMAP (n_neighbors = 30, init = random, MN_ratio = 0.1, FP_ratio = 1). Is there a best representation of the original data or are they all providing  equivalent information?"
#| label: fig-diffnldrpbmc
#| fig-pos: H
#| out-width: 100%

umap_plot_pbmc + umap_plot_pbmc_1 + umap_plot_pbmc_2 + umap_plot_pbmc_3 + umap_plot_pbmc_4 + tsne_plot_pbmc_1 + tsne_plot_pbmc + tsne_plot_pbmc_2 + phate_plot_pbmc_1 + phate_plot_pbmc_2 + phate_plot_pbmc_3 + trimap_plot_pbmc_1 + trimap_plot_pbmc_2 + trimap_plot_pbmc_3 + pacmap_plot_pbmc_1 + pacmap_plot_pbmc_2 + pacmap_plot_pbmc_3 + pacmap_plot_pbmc_4 + plot_layout(ncol = 6)
```

### 2024-03-27 paper.tex
Added figure description

@@

\includegraphics{paper_files/figure-pdf/fig-umapauthor-1.pdf}

@@

\caption{\label{fig-umapauthor}2D layout from UMAP (n\_neighbors = 30,
min\_dist = 0.3) applied for the PBMC3k dataset. Is this a best
representation of the original data?}

>>>>

\section{Introduction}\label{introduction}

@@

\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-diffnldrpbmc-1.pdf}

@@

\caption{\label{fig-diffnldrpbmc}2D layouts from different NLDR
techniques and different hyperparameter choices applied for the PBMC3k
dataset: (a) UMAP (n\_neighbors = 30, min\_dist = 0.3), (b) UMAP
(n\_neighbors = 5, min\_dist = 0.01), (c) UMAP (n\_neighbors = 15,
min\_dist = 0.99), (d) UMAP (n\_neighbors = 15, min\_dist = 0.03), (e)
UMAP (n\_neighbors = 84, min\_dist = 0.99), (f) tSNE (perplexity = 5),
(g) tSNE (perplexity = 51), (h) tSNE (perplexity = 92), (i) PHATE (knn =
2), (j) PHATE (knn = 5), (k) PHATE (knn = 10), (l) TriMAP (n\_inliers =
12, n\_outliers = 4, n\_random = 3), (m) TriMAP (n\_inliers = 30,
n\_outliers = 4, n\_random = 10), (n) TriMAP (n\_inliers = 5,
n\_outliers = 2, n\_random = 2), (p) PaCMAP (n\_neighbors = 10, init =
pca, MN\_ratio = 0.5, FP\_ratio = 2), (q) PaCMAP (n\_neighbors = 30,
init = random, MN\_ratio = 0.9, FP\_ratio = 5), (r) PaCMAP (n\_neighbors
= 5, init = pca, MN\_ratio = 0.1, FP\_ratio = 1), and (s) PaCMAP
(n\_neighbors = 30, init = random, MN\_ratio = 0.1, FP\_ratio = 1). Is
there a best representation of the original data or are they all
providing equivalent information?}
